---
title: "Genetic evidence separates systemic ferritin regulation from brain-localized susceptibility to delirium"
subtitle: "Serum ferritin and delirium risk: Integrative genomic analysis identifies locus-specific signals at 19q13 with no genome-wide causal effect"
author: "Nafiseh"
date: "2025-12-20"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: hide
    df_print: paged
    theme: flatly
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.align = "center", fig.width = 7.5, fig.height = 4.8
)

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(data.table)
  library(ggplot2)
  library(knitr)
  library(kableExtra)
  library(stringr)
})

# ---------------------------
# Colour system (your palette)
# ---------------------------
COL <- list(
  ceil     = "#999ECF",
  platinum = "#EDDDEC",
  liberty  = "#60509D",
  blue     = "#3E389E",
  grape    = "#5D1451",
  lilac    = "#adbae3",
  butter   = "#FFEDA8",
  pancake  = "#D99058",
  rose     = "#E5C8CD",
  pine     = "#1D6C5C"
)

# Accent assignment (consistent throughout report)
COL_FERRITIN <- COL$blue
COL_DELIRIUM <- COL$pine
COL_NEUTRAL  <- "#333333"

# ---------------------------
# Paths (GitHub Pages-friendly)
# Assumes this Rmd lives in docs/
# ---------------------------
DOCS_DIR <- if (basename(getwd()) == "docs") getwd() else file.path(getwd(), "docs")
FIG_DIR  <- file.path(DOCS_DIR, "figures")
DATA_DIR <- file.path(DOCS_DIR, "data")
dir.create(FIG_DIR,  showWarnings = FALSE, recursive = TRUE)
dir.create(DATA_DIR, showWarnings = FALSE, recursive = TRUE)

# ---------------------------
# Helpers
# ---------------------------
fmt_p <- function(p) {
  ifelse(is.na(p), NA_character_,
         ifelse(p < 1e-3, format(p, scientific = TRUE, digits = 2), formatC(p, digits = 3, format = "f")))
}
fmt_num <- function(x, d = 3) ifelse(is.na(x), NA_character_, formatC(x, digits = d, format = "f"))
fmt_e   <- function(x, d = 2) ifelse(is.na(x), NA_character_, format(x, scientific = TRUE, digits = d))

or_ci <- function(beta, se){
  OR  <- exp(beta)
  LCI <- exp(beta - 1.96 * se)
  UCI <- exp(beta + 1.96 * se)
  c(OR = OR, LCI = LCI, UCI = UCI)
}

kbl_nice <- function(df, caption = NULL){
  kable(df, format = "html", caption = caption, escape = FALSE) |>
    kable_styling(
      full_width = FALSE,
      position = "left",
      bootstrap_options = c("striped", "hover", "condensed", "responsive")
    ) |>
    row_spec(0, background = COL$platinum, color = COL$grape)
}

# Tooltip bubble (CSS hover). Usage: tip("definition text")
tip <- function(def){
  sprintf('<span class="help">?<span class="helptext">%s</span></span>', def)
}


dl_btn <- function(label, href){
  sprintf('<a class="dl-btn" href="%s" download>%s</a>', href, label)
}

```

<style> :root{ --ceil: #999ECF; --platinum: #EDDDEC; --liberty: #60509D; --blue: #3E389E; --grape: #5D1451; --lilac: #adbae3; --butter: #FFEDA8; --pancake: #D99058; --rose: #E5C8CD; --pine: #1D6C5C; } h1, h2, h3 { color: var(--grape); } h2 { margin-top: 1.2em; border-bottom: 2px solid var(--platinum); padding-bottom: 0.25em; } h3 { margin-top: 1.0em; } .callout { background: var(--platinum); border-left: 6px solid var(--blue); padding: 12px 14px; border-radius: 14px; margin: 10px 0 16px 0; } .cards { display: grid; grid-template-columns: repeat(auto-fit, minmax(210px, 1fr)); gap: 10px; margin: 10px 0 16px 0; } .cardx{ background: white; border: 1px solid var(--platinum); border-radius: 16px; padding: 10px 12px; box-shadow: 0 6px 18px rgba(0,0,0,0.04); } .cardx .k { color: #555; font-size: 0.88em; } .cardx .v { color: var(--grape); font-size: 1.25em; font-weight: 700; } .badge{ display:inline-block; padding: 2px 10px; border-radius: 999px; font-size: 0.85em; font-weight: 600; background: var(--butter); color: #3a2d00; border: 1px solid rgba(0,0,0,0.06); } details{ background: white; border: 1px solid var(--platinum); border-radius: 16px; padding: 10px 12px; margin: 12px 0 16px 0; } summary{ cursor: pointer; font-weight: 700; color: var(--blue); } summary:hover{ color: var(--liberty); } .dl-btn{ display: inline-block; padding: 8px 12px; border-radius: 999px; background: var(--blue); color: white !important; text-decoration: none !important; font-weight: 700; font-size: 0.90em; margin: 6px 8px 6px 0; } .dl-btn:hover{ background: var(--liberty); } .help{ position: relative; display: inline-block; width: 18px; height: 18px; line-height: 18px; text-align: center; border-radius: 999px; background: var(--lilac); color: var(--grape); font-weight: 800; font-size: 0.85em; margin-left: 6px; vertical-align: text-top; } .helptext{ visibility: hidden; opacity: 0; width: 320px; background: var(--platinum); color: #1f1f1f; border: 1px solid var(--ceil); border-radius: 14px; padding: 10px 12px; position: absolute; z-index: 999; top: 24px; left: -140px; box-shadow: 0 8px 22px rgba(0,0,0,0.08); transition: opacity 0.15s ease-in-out; } .helptext{ visibility: hidden; opacity: 0; width: 320px; background: var(--platinum); color: #1f1f1f; border: 1px solid var(--ceil); border-radius: 14px; padding: 10px 12px; position: absolute; z-index: 999; top: 24px; left: -140px; box-shadow: 0 8px 22px rgba(0,0,0,0.08); transition: opacity 0.15s ease-in-out; } .help:hover .helptext{ visibility: visible; opacity: 1; text-align: justify; text-justify: inter-word; hyphens: auto;} .figcap{ margin-top: 8px; margin-bottom: 26px; font-size: 0.95em; color: #444; background: #fafafa; border: 1px solid var(--platinum); border-radius: 14px; padding: 10px 12px; } .figcap b{ color: var(--grape); } .scroll-x{overflow-x: auto; overflow-y: hidden; width: 100%; padding-bottom: 6px;  } .scroll-xy{overflow: auto; width: 100%; max-height: 420px; padding-bottom: 6px; }
</style>
  
## Authors List:

 1.  Amirhossein Saed
 2.  Mahdi Akbarzadeh
 3.  Fatemeh Gohari
 4.  Nafiseh Asadrouh
 5.  Amir Mohammad Jahromizadeh
 6.  Hani Sabaie
 7.  Maryam Zarkesh
 8.  Mehdi Hedayati
 9.  Fereidoun Azizi
 10. Maryam Sadat Daneshpour

## Data and preprocessing

### Introduction

Exposure: Ferritin \| GWAS summary statistics: [GWAS catalogue](https://www.ebi.ac.uk/gwas/studies/GCST90270865) \| Reference paper: [Butler-Laporte G et al.2023](https://pubmed.ncbi.nlm.nih.gov/36773317/)

-   Sample size: 270,794

-   Ancestry: European

Outcome: Delirium \| GWAS summary statistics: [GWAS catalogue](https://www.ebi.ac.uk/gwas/studies/GCST90473243) \| Reference paper: [UK Biobank Whole-Genome Sequencing Consortium](https://www.ebi.ac.uk/gwas/studies/GCST90473243)

-   Sample size: 458,440

-   Ancestry: European

### Data Prepration

**Ferritin GWAS:**

-   Number of total SNPs in the raw exposure file: 10,251,641
-   Number of total SNPs in the harmonized exposure file: 10,218,238
-   Number of total SNPs with invalid rsID: 0
-   Number of total SNPs with invalid alleles: 0
-   Number of total SNPs with 0.42 \< MAF \< 0.58: 0
-   Number of total SNPs with non-numeric se/b/p: 0

**Delirium GWAS:**

-   Number of total SNPs in the raw exposure file: 96,468,105
-   Number of total SNPs in the harmonized exposure file: 84,399,074
-   Number of total SNPs with invalid rsID: 0
-   Number of total SNPs with invalid alleles: 7,725,089
-   Number of total SNPs with 0.42 \< MAF \< 0.58: 67,065,071
-   Number of total SNPs with non-numeric se/b/p: 0

------------------------------------------------------------------------

## Integrated results across MR, SMR/HEIDI, and colocalisation {.tabset .tabset-fade}

### **Two-sample MR** 

```{r tab1_mr_objects}
# ============================================================
# Tab 1 — Two-sample MR
# Values sourced from your MR log (TwoSampleMR + MR-PRESSO) and scripts.
# ============================================================

mean_F <- 78.76

# --- Main MR estimates (TwoSampleMR)
mr_main <- tibble::tribble(
  ~method, ~nsnp, ~beta, ~se, ~pval,
  "IVW (multiplicative random effects)", 70L, 0.0822032, 0.0764773, 0.2824330,
  "MR-Egger",                            70L, 0.1617036, 0.1414843, 0.2570848,
  "Weighted median",                     70L, 0.0898712, 0.0974262, 0.3562921
) |>
  rowwise() |>
  mutate(
    OR  = or_ci(beta, se)["OR"],
    LCI = or_ci(beta, se)["LCI"],
    UCI = or_ci(beta, se)["UCI"]
  ) |>
  ungroup()

# --- Heterogeneity (Cochran's Q)
het <- tibble::tribble(
  ~method, ~Q, ~Q_df, ~Q_pval,
  "IVW",      96.99547, 69L, 0.0147892,
  "MR-Egger", 96.36125, 68L, 0.0134358
)

# --- MR-Egger intercept (directional pleiotropy)
pleio <- tibble::tibble(
  egger_intercept = -0.0034819,
  se = 0.0052046,
  pval = 0.5057626
)

# --- Steiger directionality
steiger <- tibble::tibble(
  snp_r2_exposure = 0.0206426,
  snp_r2_outcome  = 0.0002185,
  correct_causal_direction = TRUE,
  steiger_pval = 0
)

# --- MR-PRESSO summary
presso <- tibble::tribble(
  ~analysis, ~beta, ~se, ~pval,
  "Raw",               0.0878940, 0.0762235, 0.2527882,
  "Outlier-corrected", 0.1198282, 0.0585500, 0.0445093
) |>
  rowwise() |>
  mutate(
    OR  = or_ci(beta, se)["OR"],
    LCI = or_ci(beta, se)["LCI"],
    UCI = or_ci(beta, se)["UCI"]
  ) |>
  ungroup()

presso_global_p    <- 0.017
presso_distort_p   <- 0.705
presso_outlier_n   <- 1L  # from your narrative: a single flagged index downstream

# --- Export tidy TSVs to docs/data/ for download buttons
fwrite(as.data.table(mr_main), file.path(DATA_DIR, "tab1_mr_main.tsv"), sep = "\t")
fwrite(as.data.table(het),     file.path(DATA_DIR, "tab1_mr_heterogeneity.tsv"), sep = "\t")
fwrite(as.data.table(pleio),   file.path(DATA_DIR, "tab1_mr_pleiotropy.tsv"), sep = "\t")
fwrite(as.data.table(steiger), file.path(DATA_DIR, "tab1_mr_steiger.tsv"), sep = "\t")
fwrite(as.data.table(presso),  file.path(DATA_DIR, "tab1_mr_presso.tsv"), sep = "\t")

```

```{r tab1_mr_header, results="asis"}
ivw <- mr_main |> dplyr::filter(method == "IVW (multiplicative random effects)") |> dplyr::slice(1)

cat(sprintf('
<div class="callout">
<b>Integrated MR interpretation.</b>
The primary IVW (random-effects) estimate is <b>non-significant</b>
(OR = %.3f, 95%% CI %.3f–%.3f; p = %s).
There is evidence of heterogeneity (IVW Cochran’s Q p = %s), but no evidence of <i>directional</i> pleiotropy (MR-Egger intercept p = %s).
Steiger filtering supports the direction Ferritin → Delirium (R²<sub>x</sub> > R²<sub>y</sub>; p = %s).
MR-PRESSO detects outliers (global p = %s), but the distortion test is non-significant (p = %s), so the outlier-corrected nominal association should be treated as <b>exploratory</b>.
</div>',
  ivw$OR, ivw$LCI, ivw$UCI, fmt_p(ivw$pval),
  fmt_p(het |> dplyr::filter(method=="IVW") |> dplyr::pull(Q_pval)),
  fmt_p(pleio$pval),
  fmt_p(steiger$steiger_pval),
  fmt_p(presso_global_p),
  fmt_p(presso_distort_p)
))

cat('<div class="cards">')
cat(sprintf('<div class="cardx"><div class="k">Instruments</div><div class="v">%d SNPs</div><span class="badge">Ferritin → Delirium</span></div>',
            ivw$nsnp))
cat(sprintf('<div class="cardx"><div class="k">Instrument strength %s</div><div class="v">%s</div><div class="k">Mean F-statistic</div></div>',
            tip("F-statistic gauges instrument strength; a common rule-of-thumb is F ≥ 10. Mean F = 78.76 indicates strong instruments and low risk of weak-IV bias."),
            fmt_num(mean_F,2)))
cat(sprintf('<div class="cardx"><div class="k">Heterogeneity %s</div><div class="v">Q p = %s</div><div class="k">Cochran’s Q (IVW)</div></div>',
            tip("Cochran’s Q tests whether per-SNP ratio estimates vary more than expected by chance; significant Q suggests heterogeneity (often pleiotropy or multiple mechanisms)."),
            fmt_p(het |> dplyr::filter(method=="IVW") |> dplyr::pull(Q_pval))))
cat(sprintf('<div class="cardx"><div class="k">Directional pleiotropy %s</div><div class="v">p = %s</div><div class="k">MR-Egger intercept</div></div>',
            tip("MR-Egger intercept tests unbalanced (directional) horizontal pleiotropy; non-significant intercept suggests no consistent directional bias."),
            fmt_p(pleio$pval)))
cat(sprintf('<div class="cardx"><div class="k">Directionality %s</div><div class="v">%s</div><div class="k">Steiger p = %s</div></div>',
            tip("Steiger compares variance explained in exposure vs outcome; if R²_exposure > R²_outcome, Ferritin → Delirium is supported over reverse."),
            ifelse(steiger$correct_causal_direction, "Supported", "Uncertain"),
            fmt_p(steiger$steiger_pval)))
cat(sprintf('<div class="cardx"><div class="k">Outliers %s</div><div class="v">%d</div><div class="k">Global p = %s; Distortion p = %s</div></div>',
            tip("MR-PRESSO tests for outlier instruments (global test) and whether removing them significantly changes the estimate (distortion test)."),
            presso_outlier_n, fmt_p(presso_global_p), fmt_p(presso_distort_p)))
cat('</div>')

```
```{r tab1_mr_tables, results="asis"}
# Table S1: main (and robust) estimators
mr_tbl <- mr_main |>
  dplyr::transmute(
    Method = method,
    `N SNPs` = nsnp,
    `β (log OR)` = fmt_num(beta, 4),
    `SE` = fmt_num(se, 4),
    `OR (95% CI)` = sprintf("%.3f (%.3f–%.3f)", OR, LCI, UCI),
    `p` = fmt_p(pval)
  )

cat(dl_btn("Download: MR main table", "data/tab1_mr_main.tsv"))
cat(kbl_nice(mr_tbl, caption = "Table S1. Two-sample MR estimates for ferritin → delirium across estimators."))

# Table S2: diagnostics “scoreboard”
diag_tbl <- tibble::tibble(
  Test = c(
    paste0("Instrument strength ", tip("Mean F-statistic summarises strength; a common heuristic is F ≥ 10.")),
    paste0("Heterogeneity (IVW) ", tip("Cochran’s Q p-value tests whether ratio estimates are heterogeneous.")),
    paste0("Directional pleiotropy ", tip("MR-Egger intercept p-value tests directional pleiotropy.")),
    paste0("Directionality (Steiger) ", tip("Steiger supports Ferritin → Delirium if R²_exposure > R²_outcome.")),
    paste0("Outliers (MR-PRESSO) ", tip("Global test detects outliers; distortion test asks if estimate changes materially after removal.")),
    "MR-PRESSO outlier-corrected estimate"
  ),
  Result = c(
    sprintf("Mean F = %s", fmt_num(mean_F,2)),
    sprintf("Q = %s (df=%d), p = %s",
            fmt_num(het$Q[het$method=="IVW"],2), het$Q_df[het$method=="IVW"], fmt_p(het$Q_pval[het$method=="IVW"])),
    sprintf("Intercept = %s (SE=%s), p = %s",
            fmt_num(pleio$egger_intercept,4), fmt_num(pleio$se,4), fmt_p(pleio$pval)),
    sprintf("R²x = %s; R²y = %s; direction = %s; p = %s",
            fmt_num(steiger$snp_r2_exposure,6), fmt_num(steiger$snp_r2_outcome,6),
            ifelse(steiger$correct_causal_direction, "Ferritin → Delirium", "Reverse/uncertain"),
            fmt_p(steiger$steiger_pval)),
    sprintf("Global p = %s; Distortion p = %s; outliers = %d",
            fmt_p(presso_global_p), fmt_p(presso_distort_p), presso_outlier_n),
    {
      pc <- presso |> dplyr::filter(analysis=="Outlier-corrected") |> dplyr::slice(1)
      sprintf("β = %s (SE=%s), OR = %.3f (%.3f–%.3f), p = %s",
              fmt_num(pc$beta,4), fmt_num(pc$se,4), pc$OR, pc$LCI, pc$UCI, fmt_p(pc$pval))
    }
  )
)

cat(dl_btn("Download: MR heterogeneity", "data/tab1_mr_heterogeneity.tsv"))
cat(dl_btn("Download: MR pleiotropy", "data/tab1_mr_pleiotropy.tsv"))
cat(dl_btn("Download: MR Steiger", "data/tab1_mr_steiger.tsv"))
cat(dl_btn("Download: MR-PRESSO", "data/tab1_mr_presso.tsv"))

cat(kbl_nice(diag_tbl, caption = "Table S2. Diagnostic summary for strength, heterogeneity, pleiotropy, directionality, and outliers."))
```


```{r tab1_forest_plot}
# Forest plot uses the exact values in mr_main + MR-PRESSO outlier-corrected
forest_df <- dplyr::bind_rows(
  mr_main |> dplyr::select(method, beta, se),
  presso  |> dplyr::filter(analysis=="Outlier-corrected") |>
    dplyr::transmute(method="IVW (MR-PRESSO outlier-corrected)", beta=beta, se=se)
) |>
  dplyr::rowwise() |>
  dplyr::mutate(
    OR  = as.numeric(or_ci(beta, se)["OR"]),
    LCI = as.numeric(or_ci(beta, se)["LCI"]),
    UCI = as.numeric(or_ci(beta, se)["UCI"])
  ) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    method = factor(method, levels = c(
      "Weighted median",
      "MR-Egger",
      "IVW (MR-PRESSO outlier-corrected)",
      "IVW (multiplicative random effects)"
    ))
  )

p_forest <- ggplot(forest_df, aes(x = OR, y = method)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = COL$ceil, linewidth = 0.9) +
  geom_errorbarh(aes(xmin = LCI, xmax = UCI), height = 0.16, colour = COL_FERRITIN, linewidth = 1.0) +
  geom_point(size = 2.8, colour = COL$grape) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )

p_forest
ggsave(file.path(FIG_DIR, "Figure_MR1_forest_theme.png"), p_forest,
       width = 7.8, height = 4.2, dpi = 320, bg = "white")

```


<div class="figcap">
  <b>Figure S1. Causal estimates across MR estimators.</b>
  Estimates are directionally consistent but imprecise across methods; the primary IVW (random-effects) interval includes the null.
</div>

<details>
<summary>Diagnostics (MR)</summary>

```{r tab1_diag_images, results="asis"}

render_fig <- function(title, filename, caption){
  fp <- file.path(FIG_DIR, filename)
  if (!file.exists(fp)) {
    cat(sprintf(
      "<p><b>Missing figure:</b> <code>%s</code><br/>Expected at: <code>%s</code></p>",
      filename, fp
    ))
    return(invisible(FALSE))
  }

  # Use repo-relative path for HTML (works on GitHub Pages)
  src <- sprintf("figures/%s", filename)

  cat(sprintf("<h4>%s</h4>", title))
  cat(sprintf(
    "<img src='%s' style='width:100%%; border-radius:16px; border:1px solid %s;'/>",
    src, COL$platinum
  ))
  cat(sprintf(
    "<div class='figcap'><b>Interpretive caption.</b> %s</div>",
    caption
  ))

  invisible(TRUE)
}

# --- Exact filenames ---
fn_scatter <- "MR scatter plot (per-SNP effects).png"
fn_funnel  <- "Funnel plot (per-SNP ratio estimates).png"
fn_loo     <- "Leave-one-out influence plot.png"
fn_presso  <- "MR-PRESSO outlier map.png"

render_fig(
  "Figure S2. MR scatter plot",
  fn_scatter,
  "Scatter of per-SNP effects (ferritin vs delirium) with estimator lines; approximate alignment of slopes supports method-level consistency in direction."
)

render_fig(
  "Figure S3. Funnel plot",
  fn_funnel,
  "Funnel symmetry provides a qualitative screen for outlier/asymmetry patterns that can reflect pleiotropy or influential instruments."
)

render_fig(
  "Figure S4. Leave-one-out influence plot",
  fn_loo,
  "Stability of the IVW estimate under leave-one-out suggests the inference is not driven by a single SNP; large shifts would indicate sensitivity to specific instruments."
)

render_fig(
  "Figure S5. MR-PRESSO outlier map",
  fn_presso,
  "MR-PRESSO outlier map visualises instruments flagged as outliers; interpretation should be paired with the global outlier test and the distortion test."
)

```

</details> 

### **SMR/HEIDI**

```{r tab2_smr_load, include=FALSE}
S1_FN <- "SuppTable_S1_SMR_HEIDI_panel_summary.tsv"
S2_FN <- "SuppTable_S2_SMR_HEIDI_BF_hits_probe_level.tsv"

S1_PATH <- file.path(DATA_DIR, S1_FN)
S2_PATH <- file.path(DATA_DIR, S2_FN)

if (!file.exists(S1_PATH)) stop("Missing file: ", S1_PATH, "\nCopy it into docs/data/ and knit again.")
if (!file.exists(S2_PATH)) stop("Missing file: ", S2_PATH, "\nCopy it into docs/data/ and knit again.")

S1 <- data.table::fread(S1_PATH)
S2 <- data.table::fread(S2_PATH)

# ---- Yields (exact; from S2) ----
yields <- S2[, .(
  records = .N,
  unique_probes = data.table::uniqueN(probe_id),
  unique_genes  = data.table::uniqueN(gene)
), by = trait]

f_y <- yields[trait == "Ferritin"]
d_y <- yields[trait == "Delirium"]

# ---- Strict overlap (genes) ----
overlap_genes <- intersect(
  unique(S2[trait == "Ferritin", gene]),
  unique(S2[trait == "Delirium", gene])
)

# ---- Convenience subsets ----
S2_ferr <- S2[trait == "Ferritin"]
S2_del  <- S2[trait == "Delirium"]

# ---- Heatmap data: tissue × omics yields (sum passing probe records) ----
heat_df <- S1[, .(n_pass_total = sum(n_pass_probes, na.rm = TRUE)), by = .(trait, tissue, omics)]
heat_df[, tissue := factor(tissue, levels = c("Blood","Brain","Pituitary","Liver","Spleen","KidneyCortex"))]
heat_df[, omics  := factor(omics,  levels = c("eQTL","mQTL","sQTL","pQTL"))]

# ---- Ferritin: top panels by passing probes (from S1) ----
top_panels_ferr <- S1[trait == "Ferritin"][order(-n_pass_probes)][1:12]
top_panels_ferr[, panel_label := paste(tissue, omics, qtl_name, sep = " | ")]

# ---- Ferritin: top genes by best (min) p_SMR per gene (from S2) ----
top_genes_ferr <- S2_ferr[, .(min_p_smr = min(p_smr, na.rm = TRUE)), by = gene][order(min_p_smr)][1:15]
top_genes_ferr[, neglog10 := -log10(min_p_smr)]
top_genes_ferr[, gene := factor(gene, levels = rev(gene))]

# ---- Delirium: best per gene (mini-bar; from S2) ----
del_gene_best <- S2_del[, .(min_p_smr = min(p_smr, na.rm = TRUE)), by = gene][order(min_p_smr)]
del_gene_best[, neglog10 := -log10(min_p_smr)]
del_gene_best[, gene := factor(gene, levels = gene)]

# ---- TABLES ----

# Table 2-1: Delirium retained hits (all rows)
tbl_delir_hits <- S2_del[order(p_smr)][, .(
  gene, tissue, omics, qtl_name, probe_id,
  chr, pos, top_snp,
  b_smr, se_smr, p_smr,
  p_heidi, nsnp_heidi,
  p_bonf_panel, M_panel
)]

# Table 2-2: Strict overlap (APOE only; trait-specific rows)
tbl_overlap <- S2[gene %chin% overlap_genes][order(trait, p_smr)][, .(
  trait, gene, tissue, omics, qtl_name, probe_id,
  chr, pos, top_snp,
  b_smr, se_smr, p_smr,
  p_heidi, nsnp_heidi,
  p_bonf_panel, M_panel
)]

# Table 2-3: Tissue × omics yields (wide)
tbl_yield_matrix <- data.table::dcast(
  heat_df, tissue + omics ~ trait,
  value.var = "n_pass_total", fill = 0
)[order(tissue, omics)]

# Table 2-4: Full S1
tbl_S1_full <- S1[order(trait, tissue, omics, qtl_name)]

# Table 2-5: Full S2
tbl_S2_full <- S2[order(trait, p_smr)]

# Table 2-6: Ferritin top hits (probe-level preview; smallest p_SMR)
tbl_ferr_top_hits <- S2_ferr[order(p_smr)][1:15, .(
  gene, tissue, omics, qtl_name, probe_id,
  chr, pos, top_snp,
  b_smr, se_smr, p_smr,
  p_heidi, nsnp_heidi
)]
```

```{r tab2_smr_header, results="asis"}
# Panels with ≥1 passing hit (from S1)
n_panels_ferr_any <- S1[trait=="Ferritin", sum(n_pass_probes > 0, na.rm=TRUE)]
n_panels_del_any  <- S1[trait=="Delirium", sum(n_pass_probes > 0, na.rm=TRUE)]

# Callout narrative (compact)
cat(sprintf(
  '<div class="callout">
  <b>SMR/HEIDI highlights.</b>
  Across 43 cis-QTL panels per trait (tissues: Blood, Brain, Pituitary, Liver, Spleen, KidneyCortex; omics: eQTL/mQTL/sQTL/pQTL),
  Ferritin shows a broad cis-mediated architecture, whereas Delirium retains sparse signals concentrated in a few panels.
  Under strict Bonferroni+HEIDI (panel-wise Bonferroni + p<sub>HEIDI</sub> ≥ 0.01 + n<sub>SNP,HEIDI</sub> ≥ 3),
  Ferritin retains <b>%d</b> probe–panel records (<b>%d</b> unique probes; <b>%d</b> genes),
  Delirium retains <b>%d</b> probe–panel records (<b>%d</b> unique probes; <b>%d</b> genes),
  and the strict cross-trait intersection is <b>%s</b>.
  </div>',
  f_y$records, f_y$unique_probes, f_y$unique_genes,
  d_y$records, d_y$unique_probes, d_y$unique_genes,
  ifelse(length(overlap_genes)==0, "none", paste(overlap_genes, collapse=", "))
))

# Cards
cat('<div class="cards">')

cat(sprintf(
  '<div class="cardx"><div class="k">Ferritin retained</div><div class="v">%d records</div><div class="k">%d probes • %d genes</div></div>',
  f_y$records, f_y$unique_probes, f_y$unique_genes
))

cat(sprintf(
  '<div class="cardx"><div class="k">Delirium retained</div><div class="v">%d records</div><div class="k">%d probes • %d genes</div></div>',
  d_y$records, d_y$unique_probes, d_y$unique_genes
))

cat(sprintf(
  '<div class="cardx"><div class="k">Strict overlap</div><div class="v">%s</div><div class="k">Shared genes (BF+HEIDI)</div></div>',
  ifelse(length(overlap_genes)==0, "0", paste(overlap_genes, collapse=", "))
))

cat(sprintf(
  '<div class="cardx"><div class="k">Ferritin panels w/ hits</div><div class="v">%d / 43</div><div class="k">Any BF+HEIDI pass</div></div>',
  n_panels_ferr_any
))

cat(sprintf(
  '<div class="cardx"><div class="k">Delirium panels w/ hits</div><div class="v">%d / 43</div><div class="k">Any BF+HEIDI pass</div></div>',
  n_panels_del_any
))

cat('</div>')

# Download buttons for the two TSVs
cat('<div style="margin-top:10px; margin-bottom:14px;">')
cat(dl_btn("Download: Table S3 (panel summary)", paste0("data/", S1_FN)))
cat(" ")
cat(dl_btn("Download: Table S4 (BF+HEIDI probe-level hits)", paste0("data/", S2_FN)))
cat('</div>')
```


```{r tab2_fig_heatmap}
p_heat <- ggplot(heat_df, aes(x = omics, y = tissue, fill = n_pass_total)) +
  geom_tile(color = COL$platinum, linewidth = 0.6) +
  facet_wrap(~trait, nrow = 1) +
  scale_fill_gradient(low = COL$platinum, high = COL_FERRITIN) +
  labs(x = NULL, y = NULL, fill = "# passing probes") +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    strip.text = element_text(color = COL$grape, face = "bold")
  )

p_heat
ggsave(file.path(FIG_DIR, "Figure_SMR_heatmap_tissue_omics.png"), p_heat,
       width = 9.4, height = 3.8, dpi = 320, bg = "white")

```

<div class="figcap"><b>Figure S6.</b> Tissue × omics yield heatmap (sum of BF+HEIDI passing probes across panels), contrasting Ferritin’s broad cis-mediated signal landscape with Delirium’s sparse retained signals.</div> 

```{r tab2_fig_top_panels_ferr}
p_panels <- ggplot(top_panels_ferr, aes(x = reorder(panel_label, n_pass_probes), y = n_pass_probes)) +
  geom_col(fill = COL_FERRITIN, alpha = 0.92) +
  coord_flip() +
  labs(x = NULL, y = "# passing probes") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank())

p_panels
ggsave(file.path(FIG_DIR, "Figure_SMR_Ferritin_top_panels.png"), p_panels,
       width = 9.2, height = 4.6, dpi = 320, bg = "white")

```

<div class="figcap"><b>Figure S7.</b> Highest-yield Ferritin panels by number of BF+HEIDI passing probes, highlighting where cis-mediated associations are most concentrated.</div>

```{r tab2_fig_top_genes_ferr}
p_genes <- ggplot(top_genes_ferr, aes(x = gene, y = neglog10)) +
  geom_segment(aes(xend = gene, y = 0, yend = neglog10), color = COL$ceil, linewidth = 1.1) +
  geom_point(size = 3.1, color = COL$grape) +
  coord_flip() +
  labs(x = NULL, y = expression(-log[10](min(p[SMR])))) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank())

p_genes
ggsave(file.path(FIG_DIR, "Figure_SMR_Ferritin_top_genes_lollipop.png"), p_genes,
       width = 8.2, height = 4.6, dpi = 320, bg = "white")

```

<div class="figcap"><b>Figure S8.</b> Top Ferritin genes ranked by the most significant SMR association per gene (minimum p<sub>SMR</sub> across BF+HEIDI passing records).</div> 

```{r tab2_fig_delir_locus}
p_del <- ggplot(del_gene_best, aes(x = gene, y = neglog10)) +
  geom_col(fill = COL_DELIRIUM, alpha = 0.90) +
  labs(x = NULL, y = expression(-log[10](min(p[SMR])))) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank())

p_del
ggsave(file.path(FIG_DIR, "Figure_SMR_Delirium_locus_minibar.png"), p_del,
       width = 6.6, height = 3.6, dpi = 320, bg = "white")

```

<div class="figcap"><b>Figure S9.</b> Delirium retained genes (best p<sub>SMR</sub> per gene) show a sparse profile dominated by the 19q13 locus signals.</div>

```{r tab2_tables_main, results="asis"}
# ---------
# Table S5: Delirium retained hits (all rows)
# ---------
t1 <- data.table::copy(tbl_delir_hits)
t1[, b_smr := fmt_num(as.numeric(b_smr), 6)]
t1[, se_smr := fmt_num(as.numeric(se_smr), 6)]
t1[, p_smr := fmt_e(as.numeric(p_smr), 2)]
t1[, p_heidi := fmt_p(as.numeric(p_heidi))]
t1[, p_bonf_panel := fmt_e(as.numeric(p_bonf_panel), 2)]
cat('<div class="scroll-x">')
cat(kbl_nice(t1, caption = "Table S5. Delirium BF+HEIDI-retained SMR hits (all passing records)."))
cat('</div>')
# ---------
# Table S6: Strict cross-trait overlap (shared genes; APOE only here)
# ---------
t2 <- data.table::copy(tbl_overlap)
t2[, b_smr := fmt_num(as.numeric(b_smr), 6)]
t2[, se_smr := fmt_num(as.numeric(se_smr), 6)]
t2[, p_smr := fmt_e(as.numeric(p_smr), 2)]
t2[, p_heidi := fmt_p(as.numeric(p_heidi))]
t2[, p_bonf_panel := fmt_e(as.numeric(p_bonf_panel), 2)]
cat('<div class="scroll-x">')
cat(kbl_nice(t2, caption = "Table S6. Strict cross-trait overlap under BF+HEIDI (shared genes)."))
cat('</div>')
# ---------
# Table S7: Tissue × omics yields (aggregated)
# ---------
t3 <- data.table::copy(tbl_yield_matrix)
cat(kbl_nice(t3, caption = "Table S7. Sum of BF+HEIDI passing probes by tissue × omics (aggregated across panels)."))
```

<details>
<summary>Table S8: panel-level summary (all 86 panels)</summary>

```{r tab2_table_S1_full, results="asis"}
tS1 <- data.table::copy(tbl_S1_full)
tS1[, p_bonf_panel := fmt_e(as.numeric(p_bonf_panel), 2)]
tS1[, min_p_smr_pass := fmt_e(as.numeric(min_p_smr_pass), 2)]
cat('<div class="scroll-xy">')
cat(kbl_nice(tS1, caption = "Table S8 — SMR/HEIDI panel summary."))
cat('</div>')
```

</details> 

<details> 
<summary>Table S9: BF+HEIDI passing hits (probe level)</summary>

```{r tab2_table_S2_full, results="asis"}
tS2 <- data.table::copy(tbl_S2_full)
tS2[, b_smr := fmt_num(as.numeric(b_smr), 6)]
tS2[, se_smr := fmt_num(as.numeric(se_smr), 6)]
tS2[, p_smr := fmt_e(as.numeric(p_smr), 2)]
tS2[, p_heidi := fmt_p(as.numeric(p_heidi))]
tS2[, p_bonf_panel := fmt_e(as.numeric(p_bonf_panel), 2)]
cat('<div class="scroll-xy">')
cat(kbl_nice(tS2, caption = "Table S9 — Bonferroni+HEIDI retained SMR hits."))
cat('</div>')
```

</details> 

<details> 
<summary>Table S10 (preview): Ferritin top BF+HEIDI passing records (smallest p<sub>SMR</sub>)</summary>

```{r tab2_table_6_full, results="asis"}
t6 <- data.table::copy(tbl_ferr_top_hits)
t6[, b_smr := fmt_num(as.numeric(b_smr), 6)]
t6[, se_smr := fmt_num(as.numeric(se_smr), 6)]
t6[, p_smr := fmt_e(as.numeric(p_smr), 2)]
t6[, p_heidi := fmt_p(as.numeric(p_heidi))]
cat('<div class="scroll-xy">')
cat(kbl_nice(t6, caption = "Table S10 (preview) - Top Ferritin BF+HEIDI passing records ranked by p_SMR."))
cat('</div>')

``` 

</details>

### **Coloc & SuSiE**

<div class="callout">
  <b>Colocalisation</b> evaluates whether two association signals within a region are consistent with a
  <b>single shared causal variant</b> (high <b>PP.H4</b>) or <b>distinct variants</b> (high <b>PP.H3</b>)
  `r tip("PP.H4 is the posterior probability that both traits share one causal variant in the tested window; PP.H3 indicates both traits have signals but are driven by different causal variants.")`.
</div>

```{r tab3_coloc_load_parse, include=FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(stringr)
  library(ggplot2)
  library(knitr)
  library(kableExtra)
})

# ---------------------------
# Required inputs (must live in docs/data/)
# ---------------------------
RUNLIST_MD <- file.path(DATA_DIR, "coloc_runlist_available_panels.md")
SUSIE_MD   <- file.path(DATA_DIR, "Susie; process and results.md")

if (!file.exists(RUNLIST_MD)) stop("Missing: ", RUNLIST_MD)
if (!file.exists(SUSIE_MD))   stop("Missing: ", SUSIE_MD)

run_lines <- readLines(RUNLIST_MD, warn = FALSE, encoding = "UTF-8")
sus_lines <- readLines(SUSIE_MD,   warn = FALSE, encoding = "UTF-8")

run_txt <- paste(run_lines, collapse = "\n")
sus_txt <- paste(sus_lines, collapse = "\n")

# ---------------------------
# Helpers
# ---------------------------

# Unicode sci-notation -> numeric
# Examples handled:
#   6.13×10⁻¹⁹⁸  -> 6.13e-198
#   1.28e−71      -> 1.28e-71   (unicode minus)
#   ~1.00         -> 1.00
sci_unicode_to_numeric <- function(x){
  if (is.na(x) || !nzchar(x)) return(NA_real_)
  x <- trimws(as.character(x))
  x <- gsub(",", "", x, fixed = TRUE)
  x <- gsub("≈", "", x, fixed = TRUE)
  x <- gsub("~", "", x, fixed = TRUE)

  # normalize unicode minus/dashes
  x <- gsub("\u2212", "-", x, fixed = TRUE)  # minus
  x <- gsub("\u2013", "-", x, fixed = TRUE)  # en dash
  x <- gsub("\u2014", "-", x, fixed = TRUE)  # em dash

  # ×10^ -> e
  x <- gsub("×10", "e", x, fixed = TRUE)
  x <- gsub("x10", "e", x, fixed = TRUE)

  # superscripts
  x <- gsub("⁻", "-", x, fixed = TRUE)
  sup <- c("⁰","¹","²","³","⁴","⁵","⁶","⁷","⁸","⁹")
  for (i in seq_along(sup)) x <- gsub(sup[i], as.character(i-1), x, fixed = TRUE)

  x <- gsub("%", "", x, fixed = TRUE)

  # strip anything not part of numeric scientific notation (SAFE char class)
  x2 <- gsub("[^0-9eE+\\.\\-]", "", x, perl = TRUE)
  suppressWarnings(as.numeric(x2))
}

grab1 <- function(pattern, text, default = NA_character_){
  m <- stringr::str_match(text, pattern)
  if (length(m) == 0 || nrow(m) == 0) return(default)
  if (ncol(m) < 2) {
    if (is.na(m[1,1])) default else m[1,1]
  } else {
    if (is.na(m[1,2])) default else m[1,2]
  }
}
grab1_num <- function(pattern, text, default = NA_real_){
  x <- grab1(pattern, text, default = NA_character_)
  sci_unicode_to_numeric(x)
}
grab1_int_commas <- function(pattern, text, default = NA_integer_){
  x <- grab1(pattern, text, default = NA_character_)
  if (is.na(x) || !nzchar(x)) return(default)
  x <- gsub(",", "", x, fixed = TRUE)
  suppressWarnings(as.integer(x))
}

# Extract a text block between two markers (regex, dotall)
block_between <- function(text, start_pat, end_pat = NULL){
  if (is.null(end_pat)) {
    out <- stringr::str_extract(text, stringr::regex(paste0("(?s)", start_pat, ".*\\Z")))
  } else {
    out <- stringr::str_extract(text, stringr::regex(paste0("(?s)", start_pat, ".*?(?=", end_pat, ")")))
  }
  if (is.na(out) || !nzchar(out)) NA_character_ else out
}

# Parse a markdown pipe table (| a | b |) into a data.frame
parse_md_pipe_table <- function(tbl_lines){
  tbl_lines <- tbl_lines[grepl("^\\|", tbl_lines)]
  if (length(tbl_lines) < 3) return(NULL)

  split_row <- function(x){
    parts <- strsplit(gsub("^\\|\\s*|\\s*\\|\\s*$", "", x), "\\|", fixed = FALSE)[[1]]
    trimws(parts)
  }
  header <- split_row(tbl_lines[1])
  rows   <- lapply(tbl_lines[-c(1,2)], split_row) # drop separator row

  rows <- lapply(rows, function(r){
    if (length(r) < length(header)) c(r, rep(NA_character_, length(header)-length(r))) else r[seq_along(header)]
  })

  df <- as.data.frame(do.call(rbind, rows), stringsAsFactors = FALSE)
  names(df) <- header
  df
}

# Find and parse the first pipe-table after a line matching `pattern`
table_after <- function(lines, pattern){
  i <- which(grepl(pattern, lines))[1]
  if (is.na(i)) return(NULL)
  j <- i + 1
  while (j <= length(lines) && trimws(lines[j]) == "") j <- j + 1
  tbl <- character(0)
  while (j <= length(lines) && grepl("^\\|", lines[j])) {
    tbl <- c(tbl, lines[j])
    j <- j + 1
  }
  parse_md_pipe_table(tbl)
}

# Best-effort numeric conversion for df columns that look numeric
numify_df <- function(df){
  if (is.null(df)) return(df)
  for (nm in names(df)) {
    x <- df[[nm]]
    x2 <- gsub(",", "", x, fixed = TRUE)
    ok <- grepl("^[-+]?[0-9]*\\.?[0-9]+([eE][-+]?[0-9]+)?$", x2) | is.na(x2) | x2=="" 
    if (all(ok)) df[[nm]] <- suppressWarnings(as.numeric(x2))
  }
  df
}

# ---------------------------
# 1) coloc.abf key results (from runlist md)
# ---------------------------

# Ferritin × APOE pQTL (primary)
sec_f_apoe <- block_between(
  run_txt,
  "### -> Ferritin-pQTL",
  "(\\n### -> Ferritin-|\\nDelirium-pQTL|\\Z)"
)
f_apoe_nsnps <- grab1_int_commas("Overlapping SNPs tested:\\s*([0-9,]+)", sec_f_apoe)
f_apoe_h0 <- grab1_num("PP\\.H0[^:]*:\\s*([0-9][^\\s\\*\\)]+)", sec_f_apoe)
f_apoe_h1 <- grab1_num("PP\\.H1[^:]*:\\s*([0-9][^\\s\\*\\)]+)", sec_f_apoe)
f_apoe_h2 <- grab1_num("PP\\.H2[^:]*:\\s*([0-9][^\\s\\*\\)]+)", sec_f_apoe)
f_apoe_h3 <- grab1_num("PP\\.H3[^:]*:\\s*([0-9][^\\s\\*\\)]+)", sec_f_apoe)
f_apoe_h4 <- grab1_num("PP\\.H4[^:]*:\\s*([0-9][^\\s\\*\\)]+)", sec_f_apoe)

# Ferritin × SLC11A2 eQTLGen row (percentages table row)
slc_line <- run_lines[grepl("^\\|\\s*SLC11A2\\s*\\|\\s*eQTLGen", run_lines)][1]
slc_vals <- if (!is.na(slc_line)) strsplit(gsub("^\\|\\s*|\\s*\\|\\s*$", "", slc_line), "\\|")[[1]] else rep(NA_character_, 8)
slc_vals <- trimws(slc_vals)
# expected: Gene | Panel | nsnps | H0% | H1% | H2% | H3% | H4%
slc_nsnps <- suppressWarnings(as.integer(slc_vals[3]))
slc_h0 <- suppressWarnings(as.numeric(gsub("%","", slc_vals[4], fixed=TRUE)) / 100)
slc_h1 <- suppressWarnings(as.numeric(gsub("%","", slc_vals[5], fixed=TRUE)) / 100)
slc_h2 <- suppressWarnings(as.numeric(gsub("%","", slc_vals[6], fixed=TRUE)) / 100)
slc_h3 <- suppressWarnings(as.numeric(gsub("%","", slc_vals[7], fixed=TRUE)) / 100)
slc_h4 <- suppressWarnings(as.numeric(gsub("%","", slc_vals[8], fixed=TRUE)) / 100)

# Ferritin × TF Liver sQTL (abf prose)
sec_tf <- block_between(run_txt, "#### 🧬For TF Gene🧬", "(\\n### ->|\\nDelirium-pQTL|\\Z)")
tf_nsnps <- grab1_int_commas("nsnps\\s*=\\s*(\\d+)", sec_tf)
tf_h0 <- grab1_num("PP\\.H0\\.abf\\s*=\\s*([0-9][^\\s\\)]+)", sec_tf)
tf_h1 <- grab1_num("PP\\.H1\\.abf\\s*=\\s*([0-9][^\\s\\)]+)", sec_tf)
tf_h2 <- grab1_num("PP\\.H2\\.abf\\s*=\\s*([0-9][^\\s\\)]+)", sec_tf)
tf_h3 <- grab1_num("PP\\.H3\\.abf\\s*=\\s*([0-9][^\\s\\)]+)", sec_tf)
tf_h4 <- grab1_num("PP\\.H4\\.abf\\s*=\\s*([0-9][^\\s\\)]+)", sec_tf)

# Delirium × APOE pQTL (primary bullets)
sec_d_apoe <- block_between(run_txt, "Delirium-pQTL\\s*\\n#### 🧬For APOE Gene🧬", "(\\nSecondary panels coloc:|\\Z)")
d_apoe_nsnps <- grab1_int_commas("SNPs tested:\\*\\*\\s*([0-9,]+)\\*\\*", sec_d_apoe)
d_apoe_h0 <- grab1_num("PP\\.H0[^:]*:\\*\\*\\s*([0-9][^\\s\\*\\)]+)", sec_d_apoe)
d_apoe_h1 <- grab1_num("PP\\.H1[^:]*:\\*\\*\\s*([0-9][^\\s\\*\\)]+)", sec_d_apoe)
d_apoe_h2 <- grab1_num("PP\\.H2[^:]*:\\*\\*\\s*([0-9][^\\s\\*\\)]+)", sec_d_apoe)
d_apoe_h3 <- grab1_num("PP\\.H3[^:]*:\\*\\*\\s*~?([0-9][^\\s\\*\\)]+)", sec_d_apoe)
d_apoe_h4 <- grab1_num("PP\\.H4[^:]*:\\*\\*\\s*([0-9][^\\s\\*\\)]+)", sec_d_apoe)

# Delirium × APOE sQTL proxies (Brain Cortex and BA9) — optional but we will display if found
sec_d_apoe_cortex <- block_between(run_txt, "Delirium × GTEx v10 sQTL \\(Brain Cortex\\) for APOE", "(\\n2\\.|\\Z)")
dc_nsnps <- grab1_int_commas("SNPs used:\\*\\*\\s*(\\d+)\\*\\*", sec_d_apoe_cortex)
dc_h0 <- grab1_num("PP\\.H0\\s*=\\s*([0-9][^\\s\\*\\)]+)", sec_d_apoe_cortex)
dc_h1 <- grab1_num("PP\\.H1\\s*=\\s*([0-9][^\\s\\*\\)]+)", sec_d_apoe_cortex)
dc_h2 <- grab1_num("PP\\.H2\\s*=\\s*([0-9][^\\s\\*\\)]+)", sec_d_apoe_cortex)
dc_h3 <- grab1_num("PP\\.H3\\s*=\\s*([0-9][^\\s\\*\\)]+)", sec_d_apoe_cortex)
dc_h4 <- grab1_num("PP\\.H4\\s*=\\s*([0-9][^\\s\\*\\)]+)", sec_d_apoe_cortex)

sec_d_apoe_ba9 <- block_between(run_txt, "Delirium × GTEx v10 sQTL \\(Brain BA9\\) for APOE", "(\\n3\\.|\\Z)")
db_nsnps <- grab1_int_commas("SNPs overlapping:\\*\\*\\s*(\\d+)\\*\\*", sec_d_apoe_ba9)
db_h0 <- grab1_num("PP\\.H0[^:]*:\\s*([0-9][^\\s\\*\\)]+)", sec_d_apoe_ba9)
db_h1 <- grab1_num("PP\\.H1[^:]*:\\s*([0-9][^\\s\\*\\)]+)", sec_d_apoe_ba9)
db_h2 <- grab1_num("PP\\.H2[^:]*:\\s*([0-9][^\\s\\*\\)]+)", sec_d_apoe_ba9)
db_h3 <- grab1_num("PP\\.H3[^:]*:\\s*~?([0-9][^\\s\\*\\)]+)", sec_d_apoe_ba9)
db_h4 <- grab1_num("PP\\.H4[^:]*:\\s*([0-9][^\\s\\*\\)]+)", sec_d_apoe_ba9)

# Delirium × CEACAM19 Cortex eQTL tables (already in runlist as pipe tables)
ce_sum   <- numify_df(table_after(run_lines, "Delirium × CEACAM19 .*— coloc summary"))
ce_top   <- numify_df(table_after(run_lines, "Top SNPs by PP\\.H4 \\(Delirium × CEACAM19, Cortex v8\\)"))
ce_lead  <- numify_df(table_after(run_lines, "Lead SNP details \\(aligned by chr:pos key\\)"))

# If summary exists, pick key numbers
ce_nsnps <- if (!is.null(ce_sum) && "nsnps" %in% names(ce_sum)) ce_sum$nsnps[1] else NA_real_
ce_h0 <- if (!is.null(ce_sum)) ce_sum[[grep("PP\\.H0", names(ce_sum), value=TRUE)[1]]][1] else NA_real_
ce_h1 <- if (!is.null(ce_sum)) ce_sum[[grep("PP\\.H1", names(ce_sum), value=TRUE)[1]]][1] else NA_real_
ce_h2 <- if (!is.null(ce_sum)) ce_sum[[grep("PP\\.H2", names(ce_sum), value=TRUE)[1]]][1] else NA_real_
ce_h3 <- if (!is.null(ce_sum)) ce_sum[[grep("PP\\.H3", names(ce_sum), value=TRUE)[1]]][1] else NA_real_
ce_h4 <- if (!is.null(ce_sum)) ce_sum[[grep("PP\\.H4", names(ce_sum), value=TRUE)[1]]][1] else NA_real_

# Build a clean "key coloc table"
df_coloc_key <- tibble::tibble(
  GWAS_trait = c("Ferritin","Ferritin","Ferritin","Delirium","Delirium","Delirium","Delirium"),
  QTL_or_proxy = c(
    "APOE pQTL (prot-a-131)",
    "SLC11A2 eQTLGen (WB)",
    "TF sQTL (Liver, GTEx v10)",
    "APOE pQTL (prot-a-131)",
    "APOE sQTL (Brain Cortex, GTEx v10)",
    "APOE sQTL (Brain BA9, GTEx v10)",
    "CEACAM19 eQTL (Cortex v8)"
  ),
  nsnps = c(f_apoe_nsnps, slc_nsnps, tf_nsnps, d_apoe_nsnps, dc_nsnps, db_nsnps, ce_nsnps),
  PP_H3 = c(f_apoe_h3, slc_h3, tf_h3, d_apoe_h3, dc_h3, db_h3, ce_h3),
  PP_H4 = c(f_apoe_h4, slc_h4, tf_h4, d_apoe_h4, dc_h4, db_h4, ce_h4)
) %>%
  mutate(
    inference = case_when(
      is.na(PP_H4) ~ "—",
      PP_H4 >= 0.80 ~ "Shared (H4 high)",
      PP_H3 >= 0.80 ~ "Distinct (H3 high)",
      TRUE ~ "Ambiguous / mixed"
    )
  )

# ---------------------------
# 2) SuSiE results (from susie md) — parse tables where available
# ---------------------------

# Ferritin ↔ APOE pQTL: coloc.susie summary table
susie_pairs <- numify_df(table_after(sus_lines, "coloc\\.susie summary \\(sorted by PP\\.H4\\)"))
apoe_pqtl_comp <- numify_df(table_after(sus_lines, "APOE pQTL components"))
ferr_comp     <- numify_df(table_after(sus_lines, "Ferritin components"))

# Delirium ↔ APOE pQTL summary (credible sets + leads)
sec_sus_del_apoe <- block_between(sus_txt, "SuSiE–coloc\\s*run for\\s*\\*\\*Delirium \\(GWAS\\)\\*\\* vs \\*\\*APOE pQTL\\*\\*", "(\\n### 2\\.|\\n### 6\\.|\\n### 7\\.|\\Z)")

del_cs_n  <- grab1_int_commas("Delirium:\\*\\*\\s*(\\d+)\\*\\* credible sets", sec_sus_del_apoe)
pqtl_cs_n <- grab1_int_commas("APOE pQTL:\\*\\*\\s*(\\d+)\\*\\* credible sets", sec_sus_del_apoe)

del_leads <- grab1("Delirium:.*?Leads:\\s*([^\\.\\n]+)", sec_sus_del_apoe, default = "")
pqtl_leads <- grab1("APOE pQTL:.*?Leads:\\s*([^\\.\\n]+)", sec_sus_del_apoe, default = "")

max_pp_shared <- grab1_num("max\\s*PP_shared[^0-9]*([0-9][^\\s\\)]+)", sus_txt)

df_sus_del_apoe <- tibble::tibble(
  comparison = "Delirium GWAS × APOE pQTL (SuSiE)",
  delirium_CS = del_cs_n,
  delirium_leads = del_leads,
  pqtl_CS = pqtl_cs_n,
  pqtl_leads = pqtl_leads,
  max_PP_shared = max_pp_shared
)

# ---------------------------
# 3) Figures
# ---------------------------

# PP.H4 bar plot across key tests
df_plot <- df_coloc_key %>%
  mutate(
    PP_H4_plot = ifelse(is.na(PP_H4), NA_real_, PP_H4),
    label = paste0(GWAS_trait, " × ", QTL_or_proxy),
    fill_col = ifelse(GWAS_trait=="Ferritin", COL_FERRITIN, COL_DELIRIUM)
  )

p_pph4 <- ggplot(df_plot %>% filter(!is.na(PP_H4_plot)),
                 aes(x = reorder(label, PP_H4_plot), y = PP_H4_plot, fill = GWAS_trait)) +
  geom_col(width = 0.75, alpha = 0.95) +
  coord_flip() +
  geom_hline(yintercept = 0.80, linetype = "dashed", color = COL$grape) +
  scale_fill_manual(values = c("Ferritin" = COL_FERRITIN, "Delirium" = COL_DELIRIUM)) +
  scale_y_continuous(labels = function(x) paste0(round(100*x), "%"), limits = c(0,1)) +
  labs(x = NULL, y = "PP.H4 (shared causal variant)", fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "top",
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(size = 10)
  )

# CEACAM19 per-SNP PP.H4 plot (if table exists)
p_ce_pph4 <- NULL
if (!is.null(ce_top) && all(c("snp","PP.H4") %in% names(ce_top))) {
  p_ce_pph4 <- ggplot(ce_top, aes(x = reorder(snp, `PP.H4`), y = `PP.H4`)) +
    geom_col(fill = COL_DELIRIUM, alpha = 0.95, width = 0.7) +
    coord_flip() +
    scale_y_continuous(labels = function(x) paste0(round(100*x), "%"), limits = c(0,1)) +
    labs(x = NULL, y = "Per-SNP PP.H4", title = NULL) +
    theme_minimal(base_size = 12) +
    theme(panel.grid.minor = element_blank())
}

# SuSiE heatmap of component pair PP.H4 (Ferritin × APOE pQTL) if table exists
p_susie_heat <- NULL
if (!is.null(susie_pairs) && all(c("idx1","idx2","PP.H4.abf") %in% names(susie_pairs))) {
  p_susie_heat <- ggplot(susie_pairs, aes(x = factor(idx1), y = factor(idx2), fill = `PP.H4.abf`)) +
    geom_tile(color = COL$platinum) +
    scale_fill_gradient(low = COL$platinum, high = COL_FERRITIN) +
    labs(x = "Ferritin component (idx1)", y = "pQTL component (idx2)", fill = "PP.H4") +
    theme_minimal(base_size = 12) +
    theme(panel.grid = element_blank())
}
```

```{r tab3,  results="asis"}
# ---------------------------

# Executive KPI cards (match Tab1/2 style)

# ---------------------------

n_tests <- nrow(df_coloc_key)
n_shared <- sum(df_coloc_key$PP_H4 >= 0.80, na.rm = TRUE)
n_distinct <- sum(df_coloc_key$PP_H3 >= 0.80, na.rm = TRUE)

# Specific APOE contrast (Ferritin vs Delirium)

apoe_f_h4 <- df_coloc_key %>% filter(GWAS_trait=="Ferritin", grepl("^APOE pQTL", QTL_or_proxy)) %>% pull(PP_H4) %>% .[1]
apoe_d_h4 <- df_coloc_key %>% filter(GWAS_trait=="Delirium", grepl("^APOE pQTL", QTL_or_proxy)) %>% pull(PP_H4) %>% .[1]

cat('<div class="cards">')

cat(sprintf(
'<div class="card"><div class="kpi-label">Key tests summarised</div><div class="kpi">%d</div><div class="kpi-sub">Across APOE / iron loci + proxies</div></div>',
n_tests
))
cat(sprintf(
'<div class="card"><div class="kpi-label">High PP.H4 (≥0.80)</div><div class="kpi">%d</div><div class="kpi-sub">Evidence for shared causal variant</div></div>',
n_shared
))
cat(sprintf(
'<div class="card"><div class="kpi-label">High PP.H3 (≥0.80)</div><div class="kpi">%d</div><div class="kpi-sub">Evidence for distinct causal variants</div></div>',
n_distinct
))
cat(sprintf(
'<div class="card"><div class="kpi-label">APOE pQTL colocalisation</div><div class="kpi">%s vs %s</div><div class="kpi-sub">Ferritin PP.H4 vs Delirium PP.H4</div></div>',
ifelse(is.na(apoe_f_h4), "—", sprintf("%.3f", apoe_f_h4)),
ifelse(is.na(apoe_d_h4), "—", sprintf("%.3f", apoe_d_h4))
))

cat('</div>')

# ---------------------------

# Section A: Summary across loci

# ---------------------------

cat('<details open class="deets"><summary><b>Summary across key loci</b></summary>')

df_show <- df_coloc_key %>%
mutate(
nsnps = ifelse(is.na(nsnps), "", as.character(nsnps)),
PP_H3 = ifelse(is.na(PP_H3), "", sprintf("%.3f", PP_H3)),
PP_H4 = ifelse(is.na(PP_H4), "", sprintf("%.3f", PP_H4))
) %>%
select(GWAS_trait, QTL_or_proxy, nsnps, PP_H3, PP_H4, inference)

print(kbl_nice(
df_show,
caption = "Table S11: Coloc.abf posterior summaries (PP.H3 vs PP.H4) across key loci and proxy panels."
))

cat('</details>')

# Figure: PP.H4 bar plot

cat('<div class="spacer"></div>')
print(p_pph4)
cat('<div class="figcap"><b>Figure S10.</b> PP.H4 highlights strong locus-specific sharing for several Ferritin–QTL pairs (e.g., APOE pQTL, SLC11A2 eQTLGen, TF sQTL), while Delirium–APOE pQTL shows PP.H3≈1 with PP.H4≈0, supporting distinct causal variants in the same region.</div>')

cat('<hr class="sep"/>')

# ---------------------------

# Section B: Locus deep dive — APOE

# ---------------------------

cat('<details class="deets"><summary><b>Locus deep dive: APOE</b></summary>')

cat('<p class="mini-note"><b>Why this matters.</b> APOE/TOMM40/CEACAM19 is a multi-signal LD block; coloc.abf can show H3 even when a shared subset exists, so SuSiE component-level checks are informative.</p>')

# SuSiE table: Ferritin × APOE pQTL component pairing

if (!is.null(susie_pairs)) {
sp <- susie_pairs %>%
mutate(across(where(is.numeric), ~ .)) %>%
arrange(desc(`PP.H4.abf`)) %>%
head(10)

print(kbl_nice(
sp,
caption = "Table S12: SuSiE–coloc (Ferritin GWAS × APOE pQTL): top component pairs ranked by PP.H4.abf."
))
} else {
cat('<p><b>Missing table.</b> Could not find the “coloc.susie summary” pipe-table inside <code>Susie; process and results.md</code>.</p>')
}

# Optional: Heatmap

if (!is.null(p_susie_heat)) {
cat('<div class="spacer"></div>')
print(p_susie_heat)
cat('<div class="figcap"><b>Figure S11.</b> Component-pair heatmap: bright tiles mark pairs consistent with shared variant(s); off-diagonal high values would suggest cross-component sharing.</div>')
}

# Delirium × APOE pQTL SuSiE summary

if (nrow(df_sus_del_apoe) > 0) {
df_del_show <- df_sus_del_apoe %>%
mutate(
delirium_leads = ifelse(is.na(delirium_leads), "", delirium_leads),
pqtl_leads = ifelse(is.na(pqtl_leads), "", pqtl_leads),
max_PP_shared = ifelse(is.na(max_PP_shared), "", sprintf("%.3e", max_PP_shared))
)
print(kbl_nice(
df_del_show,
caption = "Table S13: SuSiE summary for Delirium GWAS × APOE pQTL: credible sets, lead SNPs, and the best observed component overlap (PP_shared)."
))
}

cat('</details>')

cat('<hr class="sep"/>')

# ---------------------------

# Section C: Locus deep dive — CEACAM19 (Delirium × Cortex eQTL)

# ---------------------------

cat('<details class="deets"><summary><b>Locus deep dive: CEACAM19</b></summary>')

if (!is.null(ce_sum)) {
print(kbl_nice(
ce_sum,
caption = "Table S14: Delirium × CEACAM19 (Cortex v8) coloc.abf summary table (as recorded in runlist)."
))
}

if (!is.null(ce_top)) {
cat('<div class="spacer"></div>')
print(kbl_nice(
ce_top,
caption = "Table S15: Top SNPs by PP.H4 for Delirium × CEACAM19 (Cortex v8)."
))
}

if (!is.null(p_ce_pph4)) {
cat('<div class="spacer"></div>')
print(p_ce_pph4)
cat('<div class="figcap"><b>Figure S12.</b> Two SNPs carry essentially all posterior mass under H4, implying a very small credible set in this window.</div>')
}

if (!is.null(ce_lead)) {
cat('<div class="spacer"></div>')
print(kbl_nice(
ce_lead,
caption = "Table S16. Lead SNP details (aligned effects) for Delirium × CEACAM19 (Cortex v8)."
))
}

cat('</details>')
```

